---
title: "Reliability Demonstration Test"
output:
  html_vignette:
    fig_width: 7
    fig_height: 5
vignette: >
  %\VignetteIndexEntry{Reliability Demonstration Test}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The design of reliability demonstration tests (RDT) is a structured process aimed at verifying that a product meets specified reliability requirements. It involves determining how many units to test, how long to test them, and under what conditions.

## Examples

Suppose a new product has a reliability requirement of 95% with a confidence level of 90%. How long should the product be tested if there are 100 samples available to test? Assume an exponential distribution with a failure rate of 0.01. 

```{r}
library(rga)
result <- rdt(sample_size = 100, target = 0.95, conf_level = 0.90, failure_rate = 0.01, dist_type = "exponential")
print(result)
```

Now suppose a product with the same reliability requirement has 50 samples available to test. Assume a Weibull distribution with Beta of 1.5 and scale of 200. How long should the product be tested?

```{r}
rdt(sample_size = 50, target = 0.95, conf_level = 0.90, dist_type = "Weibull", shape = 1.5, scale = 200, plot = TRUE)
print(result)
```
