---
title: "Reliability Growth Forecasting"
output:
  html_vignette:
    fig_width: 7
    fig_height: 6
vignette: >
  %\VignetteIndexEntry{Forecasting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ReliaGrowR)
library(WeibullR)
```

## Introduction

Reliability Growth Analysis (RGA) tracks how a system's failure rate improves
over development and testing. Fitting a model to historical interval data allows
extrapolation of the trend forward, forecasting the number of additional failures
expected in a future operating window. Combined with a fleet of units currently
at risk, the forecast drives a simulation of *which* units fail and *when*,
producing a dataset suitable for life distribution estimation.

## Reliability Growth Analysis

The Crow-AMSAA model assumes the cumulative failure count follows a power-law
process: $N(t) = \lambda t^{\beta}$. A growth parameter $\beta < 1$ indicates
that failures are becoming less frequent over time (reliability improvement).

The analysis begins by defining the interval end-times and the number of
failures observed in each interval. A ten-interval test spanning 2000 cumulative
operating units shows a clear downward trend in failures per interval:

```{r}
times <- rep(200, 10)
failures <- c(25, 13, 7, 4, 2, 1, 1, 1, 2, 2)
```

The model is fitted with `rga()`:

```{r results = "asis"}
fit <- rga(times, failures, model_type = "Piecewise NHPP")
```

The fitted reliability growth model is plotted against the observed data:

```{r results = "asis"}
plot(fit,
  main = "Reliability Growth Analysis",
  xlab = "Cumulative Time", ylab = "Cumulative Failures"
)
cat("The estimated growth parameter (beta):", round(fit$betas$log_times[2,1], 3), "\n")
```

The cumulative operating time at the end of the test is `r sum(times)` units. The
total number of observed failures is `sum(failures)` = `r sum(failures)`. The fitted
$\beta < 1$ confirms reliability is improving over the course of the test.

## Reliability Growth Forecasting

The `predict_rga()` function extrapolates the fitted model beyond the test data.
The following forecast covers three consecutive 500-unit periods immediately
following the test end at $t = 2000$:

| Period | Cumulative time span |
|:-------|:---------------------|
| 1      | 2000 – 2500          |
| 2      | 2500 – 3000          |
| 3      | 3000 – 3500          |

Equal-width periods allow direct comparison: if reliability is growing, each
successive period is expected to yield fewer failures. The cumulative time
boundaries for the forecast periods are defined and the cumulative failure
forecast is generated at those boundaries:

```{r}
boundaries <- c(2000, 2500, 3000, 3500)
fc <- predict_rga(fit, boundaries)
plot(fc,
  main = "Reliability Growth Forecast",
  xlab = "Cumulative Time", ylab = "Cumulative Failures"
)
```

The incremental failures expected in each period are obtained by taking the difference 
between the cumulative forecast at successive boundaries:

```{r results = "asis"}
n_p1 <- round(fc$cum_failures[2] - fc$cum_failures[1])
n_p2 <- round(fc$cum_failures[3] - fc$cum_failures[2])
n_p3 <- round(fc$cum_failures[4] - fc$cum_failures[3])
cat(
  "Predicted failures — Period 1:", n_p1,
  " Period 2:", n_p2,
  " Period 3:", n_p3, "\n"
)
```

Each predicted count indicates how many new failures to inject into the fleet simulation 
for that period. The predicted interval failure counts serve as the injection rate 
for the fleet simulation below.

## Stochastic Fleet Simulation

The forecasted failure counts drive a stochastic simulation of individual unit
failures. The `sim_failures()` function implements a stochastic assignment
process that mimics field failure occurrence. Units are selected for failure
using probability proportional to size (PPS) sampling: units with longer
accumulated runtimes have a proportionally higher probability of being chosen.
Failure event times are drawn from `runtime + Uniform(0, window)` and surviving
units are right-censored at `runtime + window`.

Suppose a fleet of 40 units is currently in service, with accumulated runtimes spanning
roughly 500 to 2000 operating units, which forms the at-risk population:

```{r}
set.seed(42)
runtimes <- sort(sample(500:2000, 40, replace = FALSE))
```

### Period 1 Simulation

The Period 1 simulation injects `n_p1` failures into the 40-unit fleet over a
500-unit observation window:

```{r}
result_p1 <- sim_failures(n_p1, runtimes, window = 500)
knitr::kable(head(result_p1), caption = "Simulated failures and suspensions for Period 1")
```

### Period 2 Simulation

For Period 2, surviving units from Period 1 each accumulate an additional 500 units
of operating time. Each failed unit is replaced by a new unit (starting runtime = 1),
keeping the at-risk population roughly constant. The replacement units contribute
suspension observations at the end of the window if they do not fail, anchoring
the left tail of the life distribution estimate:

```{r}
survivors_p1 <- result_p1$index[result_p1$type == "Suspension"]
n_fail_p1   <- sum(result_p1$type == "Failure")
runtimes_p2 <- c(runtimes[survivors_p1] + 500, rep(1, n_fail_p1))
result_p2 <- sim_failures(n_p2, runtimes_p2, window = 500)
```

### Period 3 Simulation

The same runtime-advance and replacement logic applies after Period 2:

```{r}
survivors_p2 <- result_p2$index[result_p2$type == "Suspension"]
n_fail_p2   <- sum(result_p2$type == "Failure")
runtimes_p3 <- c(runtimes_p2[survivors_p2] + 500, rep(1, n_fail_p2))
result_p3 <- sim_failures(n_p3, runtimes_p3, window = 500)
```

## Parametric Life Distribution Estimation

A Weibull distribution is fitted to the combined failure and suspension data from
each period. To supplement the stochastic simulation results, approximate individual
failure times are generated from the historical interval failure counts, and the
fleet runtimes are included as right-censored observations. The combined dataset includes
historical test failures, fleet runtimes, and simulated events. Each Weibull fit is a cumulative
update: Period 2 incorporates all failure events observed through Period 1 plus the
new Period 2 failures, with right-censoring at the end of the current window. This
approach mirrors standard practice, updating the life distribution estimate as new
data arrive over time.

```{r}
# Generate approximate failure times from interval counts
interval_ends <- cumsum(times)
interval_starts <- c(0, interval_ends[-length(interval_ends)])
hist_fail_times <- sort(unlist(mapply(function(start, end, n) {
  if (n > 0) runif(n, start, end) else numeric(0)
}, interval_starts, interval_ends, failures)))

# Extract events by type from each period
fail_p1 <- result_p1$runtime[result_p1$type == "Failure"]
susp_p1 <- result_p1$runtime[result_p1$type == "Suspension"]

fail_p2 <- result_p2$runtime[result_p2$type == "Failure"]
susp_p2 <- result_p2$runtime[result_p2$type == "Suspension"]

fail_p3 <- result_p3$runtime[result_p3$type == "Failure"]
susp_p3 <- result_p3$runtime[result_p3$type == "Suspension"]

# Period 1: historical + simulated failures; fleet runtimes + simulated suspensions
obj_p1 <- wblr(c(hist_fail_times, fail_p1), c(runtimes, susp_p1),
  col = "steelblue", label = "Period 1", is.plot.cb = FALSE
)
obj_p1 <- wblr.fit(obj_p1)

# Period 2: cumulative failures through P2; fleet runtimes + P2 suspensions
obj_p2 <- wblr(c(hist_fail_times, fail_p1, fail_p2), c(runtimes, susp_p2),
  col = "tomato", label = "Period 2", is.plot.cb = FALSE
)
obj_p2 <- wblr.fit(obj_p2)

# Period 3: cumulative failures through P3; fleet runtimes + P3 suspensions
obj_p3 <- wblr(c(hist_fail_times, fail_p1, fail_p2, fail_p3), c(runtimes, susp_p3),
  col = "forestgreen", label = "Period 3", is.plot.cb = FALSE
)
obj_p3 <- wblr.fit(obj_p3)
```

All three fitted lines are overlaid on a single Weibull probability plot. As
reliability grows, the fitted line shifts rightward, indicating a longer
characteristic life:

```{r}
plot.wblr(list(obj_p1, obj_p2, obj_p3),
  main = "Cumulative Weibull Fits Across Forecast Periods",
  is.plot.legend = FALSE
)
```

## Summary: Weibull Parameters Across Periods

The estimated shape ($\beta$) and scale ($\eta$) parameters from each fitted
object are compiled into a summary table:

```{r}
cum_failures <- cumsum(lengths(list(fail_p1, fail_p2, fail_p3)))

params <- data.frame(
  Period = c("Period 1", "Period 2", "Period 3"),
  Cumul_Failures = cum_failures,
  Beta = round(c(
    obj_p1$fit[[1]]$beta,
    obj_p2$fit[[1]]$beta,
    obj_p3$fit[[1]]$beta
  ), 3),
  Eta = round(c(
    obj_p1$fit[[1]]$eta,
    obj_p2$fit[[1]]$eta,
    obj_p3$fit[[1]]$eta
  ), 1)
)

knitr::kable(params,
  caption = "Cumulative Weibull parameters by forecast period",
  col.names = c("Period", "Cumul. Failures", "Shape (\u03b2)", "Scale (\u03b7)"),
  row.names = FALSE
)
```

With each update, the estimate is based on more failure events, reducing
uncertainty. The rightward shift of scale $\eta$ across periods reflects the
reduced failure intensity predicted by the growth model, units are lasting longer
as the system matures.

### Reliability Metrics

The Weibull parameters translate into two common reliability metrics: the
mean time between failures (MTBF) and the reliability at a fixed mission time.
MTBF for a Weibull distribution is $\eta \cdot \Gamma(1 + 1/\beta)$, and
reliability at time $t$ is $R(t) = \exp[-(t/\eta)^{\beta}]$.

```{r}
weibull_metrics <- function(obj, t) {
  b <- obj$fit[[1]]$beta
  e <- obj$fit[[1]]$eta
  mtbf <- e * gamma(1 + 1 / b)
  rt <- exp(-(t / e)^b)
  c(MTBF = round(mtbf, 1), Reliability = round(rt, 3))
}

t_mission <- 1500 # target mission time (operating units)

metrics <- data.frame(
  Period = c("Period 1", "Period 2", "Period 3"),
  t(sapply(list(obj_p1, obj_p2, obj_p3), weibull_metrics, t = t_mission))
)

knitr::kable(metrics,
  caption = paste0(
    "MTBF and reliability at t = ", t_mission,
    " operating units by forecast period"
  ),
  col.names = c("Period", "MTBF", paste0("R(", t_mission, ")")),
  row.names = FALSE
)
```

A rising MTBF and reliability across periods confirms that the reliability
growth signal detected in testing carries through to the fleet simulation.

## Monte Carlo Uncertainty Analysis

A single run of `sim_failures()` represents one realization of the stochastic
assignment process. Uncertainty in $\hat{\beta}$ and $\hat{\eta}$ attributable
to randomness in failure assignment is quantified by repeating all three
simulation periods from the same initial fleet state, yielding a collection of
Weibull parameter estimates whose spread measures sampling variability.

### Run the Monte Carlo Loop

Each iteration re-simulates all three periods and fits a cumulative Weibull model 
for each period. The historical failure times and fleet runtimes are included alongside 
the simulation data in each fit. Parameters are stored in a long-format data frame. 
Iterations yielding fewer than two failures in a given period are excluded from that 
period's fit by design; errors are suppressed via `tryCatch`.

```{r mc-loop, message = FALSE, warning = FALSE, results = "hide"}
set.seed(99)
n_sim <- 200
mc_list <- vector("list", n_sim)

for (i in seq_len(n_sim)) {
  tryCatch(
    {
      # Period 1
      r1 <- sim_failures(n_p1, runtimes, window = 500)
      f1 <- r1$runtime[r1$type == "Failure"]
      s1 <- r1$runtime[r1$type == "Suspension"]

      # Period 2 — survivors advance; failed units replaced by new units (runtime = 1)
      surv1 <- r1$index[r1$type == "Suspension"]
      rt2 <- c(runtimes[surv1] + 500, rep(1, sum(r1$type == "Failure")))
      r2 <- sim_failures(n_p2, rt2, window = 500)
      f2 <- r2$runtime[r2$type == "Failure"]
      s2 <- r2$runtime[r2$type == "Suspension"]

      # Period 3 — survivors advance; failed units replaced by new units (runtime = 1)
      surv2 <- r2$index[r2$type == "Suspension"]
      rt3 <- c(rt2[surv2] + 500, rep(1, sum(r2$type == "Failure")))
      r3 <- sim_failures(n_p3, rt3, window = 500)
      f3 <- r3$runtime[r3$type == "Failure"]
      s3 <- r3$runtime[r3$type == "Suspension"]

      # Cumulative failure and suspension sets including historical + fleet data
      cum_fail <- list(
        c(hist_fail_times, f1),
        c(hist_fail_times, f1, f2),
        c(hist_fail_times, f1, f2, f3)
      )
      susp_set <- list(
        c(runtimes, s1),
        c(runtimes, s2),
        c(runtimes, s3)
      )

      rows <- lapply(1:3, function(p) {
        if (length(cum_fail[[p]]) < 2) {
          return(NULL)
        }
        obj <- wblr(cum_fail[[p]], susp_set[[p]], is.plot.cb = FALSE)
        obj <- wblr.fit(obj)
        data.frame(
          sim = i,
          period = paste0("Period ", p),
          beta = obj$fit[[1]]$beta,
          eta = obj$fit[[1]]$eta
        )
      })

      mc_list[[i]] <- do.call(rbind, Filter(Negate(is.null), rows))
    },
    error = function(e) NULL
  )
}

mc_df <- do.call(rbind, Filter(Negate(is.null), mc_list))
```

### Visualize the Parameter Distributions

Histograms of $\hat{\beta}$ (shape) and $\hat{\eta}$ (scale) for each period
show how the fitted parameters vary across simulations. The dashed vertical
line marks the median; the solid vertical line marks the single-run point
estimate from the previous section. A density curve is overlaid to highlight
the distributional shape.

```{r mc-histograms, fig.height = 6}
# Single-run point estimates for reference
single_beta <- c(obj_p1$fit[[1]]$beta, obj_p2$fit[[1]]$beta, obj_p3$fit[[1]]$beta)
single_eta <- c(obj_p1$fit[[1]]$eta, obj_p2$fit[[1]]$eta, obj_p3$fit[[1]]$eta)

cols <- c("steelblue", "tomato", "forestgreen")
par(mfrow = c(2, 3), mar = c(4, 4, 3, 1))

for (p in 1:3) {
  sub <- mc_df[mc_df$period == paste0("Period ", p), ]
  b <- sub$beta[!is.na(sub$beta)]
  h <- hist(b,
    breaks = "Sturges", col = cols[p], border = "white",
    main = paste0("Period ", p, " \u2014 Shape (\u03b2)"),
    xlab = expression(hat(beta)), ylab = "Count", freq = TRUE
  )
  # density scaled to count axis
  dens <- density(b)
  scale_factor <- length(b) * diff(h$breaks[1:2])
  lines(dens$x, dens$y * scale_factor, lwd = 2)
  abline(v = median(b), lty = 2, lwd = 2)
  abline(v = single_beta[p], lty = 1, lwd = 2, col = "black")
}

for (p in 1:3) {
  sub <- mc_df[mc_df$period == paste0("Period ", p), ]
  e <- sub$eta[!is.na(sub$eta)]
  h <- hist(e,
    breaks = "Sturges", col = cols[p], border = "white",
    main = paste0("Period ", p, " \u2014 Scale (\u03b7)"),
    xlab = expression(hat(eta)), ylab = "Count", freq = TRUE
  )
  dens <- density(e)
  scale_factor <- length(e) * diff(h$breaks[1:2])
  lines(dens$x, dens$y * scale_factor, lwd = 2)
  abline(v = median(e), lty = 2, lwd = 2)
  abline(v = single_eta[p], lty = 1, lwd = 2, col = "black")
}

par(mfrow = c(1, 1))
```

The solid line (single-run estimate) and dashed line (MC median) are expected to
align closely. Any gap reflects sampling variability in that single realization.

### Monte Carlo Summary Table

Median, mean, and 95% central intervals across all valid simulations are
presented below. Skewed parameter distributions show a mean that differs
noticeably from the median.

```{r mc-table}
mc_summary <- do.call(rbind, lapply(1:3, function(p) {
  sub <- mc_df[mc_df$period == paste0("Period ", p), ]
  b <- sub$beta[!is.na(sub$beta)]
  e <- sub$eta[!is.na(sub$eta)]
  data.frame(
    Period    = paste0("Period ", p),
    Valid     = length(b),
    Pct_valid = paste0(round(100 * length(b) / n_sim), "%"),
    Beta_mean = round(mean(b), 3),
    Beta_med  = round(median(b), 3),
    Beta_lo   = round(quantile(b, 0.025), 3),
    Beta_hi   = round(quantile(b, 0.975), 3),
    Eta_mean  = round(mean(e), 1),
    Eta_med   = round(median(e), 1),
    Eta_lo    = round(quantile(e, 0.025), 1),
    Eta_hi    = round(quantile(e, 0.975), 1)
  )
}))

knitr::kable(mc_summary,
  caption = "Monte Carlo summary: mean, median and 95% central interval (200 simulations)",
  col.names = c(
    "Period", "Valid", "%",
    "Mean \u03b2", "Med. \u03b2", "2.5% \u03b2", "97.5% \u03b2",
    "Mean \u03b7", "Med. \u03b7", "2.5% \u03b7", "97.5% \u03b7"
  ),
  row.names = FALSE
)
```

Wider intervals in Period 1 reflect the small number of simulated failures; as
failures accumulate across periods, the estimates stabilize and the intervals
narrow. A rightward shift in the median $\hat{\eta}$ across periods is the
Monte Carlo analogue of the point-estimate shift seen in the single-run
summary, now expressed as a full distribution rather than a scalar.
Iterations with fewer than two failures in a period are excluded from that
period's fit. The "%" column shows how often a valid fit was obtained.

## Sensitivity Analysis

The preceding Monte Carlo analysis fixes the growth model parameters. This
section varies the growth model forecast in two ways: by scaling the predicted
failure counts and by adjusting the growth parameter $\beta$. The resulting
Weibull fits reveal how the life distribution estimate responds to changes in
the growth model forecast.

```{r sens-setup}
# Helper: run all three periods in one call; failed units are replaced by new
# units (runtime = 1), keeping the at-risk population roughly constant.
run_three_periods <- function(n_vec, runtimes, window = 500) {
  r1 <- sim_failures(n_vec[1], runtimes, window = window)
  surv1 <- r1$index[r1$type == "Suspension"]
  n_fail1 <- sum(r1$type == "Failure")
  rt2 <- c(runtimes[surv1] + window, rep(1, n_fail1))
  if (length(rt2) < 2) stop("insufficient units after Period 1")
  r2 <- sim_failures(min(n_vec[2], length(rt2) - 1), rt2, window = window)
  surv2 <- r2$index[r2$type == "Suspension"]
  n_fail2 <- sum(r2$type == "Failure")
  rt3 <- c(rt2[surv2] + window, rep(1, n_fail2))
  if (length(rt3) < 2) stop("insufficient units after Period 2")
  r3 <- sim_failures(min(n_vec[3], length(rt3) - 1), rt3, window = window)
  list(r1 = r1, r2 = r2, r3 = r3, rt2 = rt2, rt3 = rt3)
}

cols3 <- c("steelblue", "tomato", "forestgreen")
```

### Effect of Number of Failures

More observed failures provide denser information for fitting the Weibull
distribution. Four multipliers are applied to the base predicted failure counts,
spanning a wide range to expose how sample size affects parameter precision.
A larger fleet of 80 units is used here so that even the highest failure counts
do not deplete the at-risk population before Period 3.

| Label   | Multiplier | ~Period 1 failures |
|---------|-----------|-------------------|
| Base    | 1×        | `r n_p1` |
| 2×      | 2×        | ~`r round(n_p1 * 2)` |
| 4×      | 4×        | ~`r round(n_p1 * 4)` |
| 6×      | 6×        | ~`r round(n_p1 * 6)` |

```{r sens-fail-setup}
set.seed(7)
n_sim_sens <- 200
runtimes_fail <- sort(sample(500:2000, 80, replace = FALSE))

fail_scenarios <- list(
  Base  = c(n_p1, n_p2, n_p3),
  `2x`  = pmax(round(c(n_p1, n_p2, n_p3) * 2), 1),
  `4x`  = pmax(round(c(n_p1, n_p2, n_p3) * 4), 1),
  `6x`  = pmax(round(c(n_p1, n_p2, n_p3) * 6), 1)
)
# Cap at fleet size - 1 to prevent over-sampling
fail_scenarios <- lapply(fail_scenarios, function(ns) pmin(ns, length(runtimes_fail) - 1))
```

```{r sens-fail-plots, fig.height = 8, message = FALSE, warning = FALSE}
set.seed(42)
par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

for (sc_name in names(fail_scenarios)) {
  tryCatch({
    ns <- fail_scenarios[[sc_name]]
    n1 <- ns[1]; n2 <- ns[2]; n3 <- ns[3]
    rt <- runtimes_fail

    res <- run_three_periods(c(n1, n2, n3), rt)
    f1 <- res$r1$runtime[res$r1$type == "Failure"]
    s1 <- res$r1$runtime[res$r1$type == "Suspension"]
    f2 <- res$r2$runtime[res$r2$type == "Failure"]
    s2 <- res$r2$runtime[res$r2$type == "Suspension"]
    f3 <- res$r3$runtime[res$r3$type == "Failure"]
    s3 <- res$r3$runtime[res$r3$type == "Suspension"]

    o1 <- wblr(c(hist_fail_times, f1), c(runtimes_fail, s1),
      col = cols3[1], label = "Period 1", is.plot.cb = FALSE)
    o1 <- wblr.fit(o1)
    o2 <- wblr(c(hist_fail_times, f1, f2), c(runtimes_fail, s2),
      col = cols3[2], label = "Period 2", is.plot.cb = FALSE)
    o2 <- wblr.fit(o2)
    o3 <- wblr(c(hist_fail_times, f1, f2, f3), c(runtimes_fail, s3),
      col = cols3[3], label = "Period 3", is.plot.cb = FALSE)
    o3 <- wblr.fit(o3)
    plot.wblr(list(o1, o2, o3), main = paste0(sc_name, " failures (n\u2081=", n1, ")"), 
              is.plot.legend = FALSE)
  }, error = function(e) {
    plot.new()
    title(main = paste0(sc_name, "\n(insufficient survivors)"))
  })
}
par(mfrow = c(1, 1))
```

A Monte Carlo analysis of 200 iterations per failure scenario collects
$\hat{\beta}$ and $\hat{\eta}$ for the cumulative Period 3 fit:

```{r sens-fail-mc, message = FALSE, warning = FALSE, results = "hide"}
set.seed(456)
fail_mc_list <- lapply(names(fail_scenarios), function(sc_name) {
  ns <- fail_scenarios[[sc_name]]
  n1 <- ns[1]; n2 <- ns[2]; n3 <- ns[3]

  rows <- lapply(seq_len(n_sim_sens), function(i) {
    tryCatch({
      res <- run_three_periods(c(n1, n2, n3), runtimes_fail)
      f1 <- res$r1$runtime[res$r1$type == "Failure"]
      f2 <- res$r2$runtime[res$r2$type == "Failure"]
      f3 <- res$r3$runtime[res$r3$type == "Failure"]
      s3 <- res$r3$runtime[res$r3$type == "Suspension"]

      cum_f3 <- c(hist_fail_times, f1, f2, f3)
      if (length(cum_f3) < 2) return(NULL)
      obj <- wblr(cum_f3, c(runtimes_fail, s3), is.plot.cb = FALSE)
      obj <- wblr.fit(obj)
      data.frame(scenario = sc_name, beta = obj$fit[[1]]$beta, eta = obj$fit[[1]]$eta)
    }, error = function(e) NULL)
  })
  do.call(rbind, Filter(Negate(is.null), rows))
})
names(fail_mc_list) <- names(fail_scenarios)
fail_mc_df <- do.call(rbind, fail_mc_list)
fail_mc_df$scenario <- factor(fail_mc_df$scenario, levels = names(fail_scenarios))
```

```{r sens-fail-boxplot, fig.height = 5}
fail_cols <- c("#F9E79F", "#F8C471", "#EB984E", "#E74C3C")
par(mfrow = c(1, 2))
boxplot(beta ~ scenario, data = fail_mc_df,
  xlab = "Failure scenario", ylab = expression(hat(beta)),
  main = "Period 3 Shape (\u03b2) by Failure Count",
  col  = fail_cols, outline = FALSE
)
boxplot(eta ~ scenario, data = fail_mc_df,
  xlab = "Failure scenario", ylab = expression(hat(eta)),
  main = "Period 3 Scale (\u03b7) by Failure Count",
  col  = fail_cols, outline = FALSE
)
par(mfrow = c(1, 1))
```

```{r sens-fail-table}
fail_tbl <- do.call(rbind, lapply(names(fail_scenarios), function(sc) {
  sub <- fail_mc_df[fail_mc_df$scenario == sc, ]
  b <- sub$beta[!is.na(sub$beta)]
  e <- sub$eta[!is.na(sub$eta)]
  data.frame(
    Scenario = sc,
    n_total  = sum(fail_scenarios[[sc]]),
    Valid    = length(b),
    Beta_med = round(median(b), 3),
    Beta_lo  = round(quantile(b, 0.025), 3),
    Beta_hi  = round(quantile(b, 0.975), 3),
    Eta_med  = round(median(e), 1),
    Eta_lo   = round(quantile(e, 0.025), 1),
    Eta_hi   = round(quantile(e, 0.975), 1)
  )
}))

knitr::kable(fail_tbl,
  caption = "Period 3 \u03b2 and \u03b7 summary by failure-count scenario (200 simulations each)",
  col.names = c(
    "Scenario", "Total n (3 periods)", "Valid",
    "Med. \u03b2", "2.5% \u03b2", "97.5% \u03b2",
    "Med. \u03b7", "2.5% \u03b7", "97.5% \u03b7"
  ),
  row.names = FALSE
)
```

As failure counts increase, the parameter uncertainty shrinks, the 95% intervals
for both $\hat{\beta}$ and $\hat{\eta}$ narrow, and the median $\hat{\eta}$
converges toward the true characteristic life implied by the underlying runtime
distribution.

### Effect of Growth Model Parameters ($\beta$)

The shape of the reliability growth trajectory is governed entirely by the Crow-AMSAA
growth parameter $\beta$. When $\beta < 1$ failures become less frequent over time
(reliability is improving); when $\beta = 1$ the process is a homogeneous Poisson
process (no change); and when $\beta > 1$ failures increase over time (degradation).
This subsection examines how a changing $\beta$ propagates through the forecasting and
simulation pipeline to produce distinct Weibull life distributions.

Four scenarios are constructed by generating synthetic interval failure data whose
expected counts follow a power-law process with the prescribed $\beta$.  The same
cumulative-time axis (10 intervals of 200 units, ending at $t = 2000$) and the same
approximate total failure count (52) are used across all scenarios so that only the
*shape* of the growth trajectory differs. The $\beta = 1.1$ scenario sits just above 
the growth/degradation threshold and serves as the near-linear reference.

```{r beta-setup}
# Growth parameter scenarios
beta_scenarios <- c(0.3, 0.6, 1.1, 1.4)
beta_labels <- c(
  "\u03b2 = 0.3 (strong growth)",
  "\u03b2 = 0.6 (moderate growth)",
  "\u03b2 = 1.1 (slight degradation)",
  "\u03b2 = 1.4 (degradation)"
)

# Generate interval failure counts consistent with a given beta.
# lambda is chosen so that cumulative failures at t_end ≈ total.
gen_interval_failures <- function(b, total = 52, n_int = 10, t_end = 2000) {
  t  <- seq(0, t_end, length.out = n_int + 1)
  lm <- total / t_end^b
  expected <- lm * (t[-1]^b - t[-length(t)]^b)
  pmax(round(expected), 1L)
}

# Verify the generated counts
beta_failure_counts <- lapply(beta_scenarios, gen_interval_failures)
names(beta_failure_counts) <- beta_labels
```

#### Single-run Weibull overlay per scenario

For each $\beta$ scenario, the RGA model is fitted, three 500-unit periods
beyond $t = 2000$ are forecast, and the stochastic fleet simulation is run
(40 units, with replacements). Historical failure times derived from each
scenario's interval data and the fleet runtimes are included in the Weibull
fits. A rightward shift across periods indicates that the reliability growth
trend is visible in the life distribution; a leftward shift (degradation
scenario) indicates the opposite.

```{r beta-single-run, fig.height = 8, message = FALSE, warning = FALSE}
set.seed(42)
runtimes_beta <- sort(sample(500:2000, 40, replace = FALSE))

# Pre-generate historical failure times for each scenario
hist_fail_list <- lapply(beta_scenarios, function(b) {
  fc_b <- gen_interval_failures(b)
  sort(unlist(mapply(function(start, end, n) {
    if (n > 0) runif(n, start, end) else numeric(0)
  }, interval_starts, interval_ends, fc_b)))
})

par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

beta_single <- vector("list", length(beta_scenarios))
names(beta_single) <- beta_labels

for (k in seq_along(beta_scenarios)) {
  tryCatch({
    fc_b  <- gen_interval_failures(beta_scenarios[k])
    hist_fail_b <- hist_fail_list[[k]]

    fit_b <- rga(rep(200, 10), fc_b)
    bd_b  <- c(2000, 2500, 3000, 3500)
    pred_b <- predict_rga(fit_b, bd_b)
    np1 <- max(1L, round(pred_b$cum_failures[2] - pred_b$cum_failures[1]))
    np2 <- max(1L, round(pred_b$cum_failures[3] - pred_b$cum_failures[2]))
    np3 <- max(1L, round(pred_b$cum_failures[4] - pred_b$cum_failures[3]))

    res <- run_three_periods(c(np1, np2, np3), runtimes_beta)
    f1 <- res$r1$runtime[res$r1$type == "Failure"]
    s1 <- res$r1$runtime[res$r1$type == "Suspension"]
    f2 <- res$r2$runtime[res$r2$type == "Failure"]
    s2 <- res$r2$runtime[res$r2$type == "Suspension"]
    f3 <- res$r3$runtime[res$r3$type == "Failure"]
    s3 <- res$r3$runtime[res$r3$type == "Suspension"]

    o1 <- wblr(c(hist_fail_b, f1), c(runtimes_beta, s1),
      col = cols3[1], label = "Period 1", is.plot.cb = FALSE)
    o1 <- wblr.fit(o1)
    o2 <- wblr(c(hist_fail_b, f1, f2), c(runtimes_beta, s2),
      col = cols3[2], label = "Period 2", is.plot.cb = FALSE)
    o2 <- wblr.fit(o2)
    o3 <- wblr(c(hist_fail_b, f1, f2, f3), c(runtimes_beta, s3),
      col = cols3[3], label = "Period 3", is.plot.cb = FALSE)
    o3 <- wblr.fit(o3)

    beta_single[[k]] <- list(o1 = o1, o2 = o2, o3 = o3,
                              np = c(np1, np2, np3))
    plot.wblr(list(o1, o2, o3), main = beta_labels[k], is.plot.legend = FALSE)
  }, error = function(e) {
    plot.new()
    title(main = paste0(beta_labels[k], "\n(insufficient data)"))
  })
}
par(mfrow = c(1, 1))
```

When $\beta < 1$ the Weibull lines shift rightward across periods, reflecting the
growing characteristic life predicted by the model.  Once $\beta$ exceeds 1 the
lines shift leftward, indicating earlier expected failures with each successive
period. The $\beta = 1.1$ panel illustrates this reversal: even a modest exceedance
above the growth threshold produces a visible leftward progression.

#### Monte Carlo summary by $\beta$ scenario

A Monte Carlo analysis of 200 iterations per scenario collects the Period 3
cumulative $\hat{\beta}$ (Weibull shape) and $\hat{\eta}$ (Weibull scale):

```{r beta-mc, message = FALSE, warning = FALSE, results = "hide"}
set.seed(321)
n_sim_beta <- 200

beta_mc_list <- lapply(seq_along(beta_scenarios), function(k) {
  hist_fail_b <- hist_fail_list[[k]]

  fc_b <- gen_interval_failures(beta_scenarios[k])
  fit_b <- rga(rep(200, 10), fc_b)
  bd_b  <- c(2000, 2500, 3000, 3500)
  pred_b <- predict_rga(fit_b, bd_b)
  np1 <- max(1L, round(pred_b$cum_failures[2] - pred_b$cum_failures[1]))
  np2 <- max(1L, round(pred_b$cum_failures[3] - pred_b$cum_failures[2]))
  np3 <- max(1L, round(pred_b$cum_failures[4] - pred_b$cum_failures[3]))

  rows <- lapply(seq_len(n_sim_beta), function(i) {
    tryCatch({
      res <- run_three_periods(c(np1, np2, np3), runtimes_beta)
      f1  <- res$r1$runtime[res$r1$type == "Failure"]
      f2  <- res$r2$runtime[res$r2$type == "Failure"]
      f3  <- res$r3$runtime[res$r3$type == "Failure"]
      s3  <- res$r3$runtime[res$r3$type == "Suspension"]
      cum_f <- c(hist_fail_b, f1, f2, f3)
      if (length(cum_f) < 2) return(NULL)
      obj <- wblr(cum_f, c(runtimes_beta, s3), is.plot.cb = FALSE)
      obj <- wblr.fit(obj)
      data.frame(beta_rga = beta_scenarios[k],
                 label    = beta_labels[k],
                 beta_wb  = obj$fit[[1]]$beta,
                 eta      = obj$fit[[1]]$eta)
    }, error = function(e) NULL)
  })
  do.call(rbind, Filter(Negate(is.null), rows))
})
beta_mc_df <- do.call(rbind, beta_mc_list)
beta_mc_df$label <- factor(beta_mc_df$label, levels = beta_labels)
```

```{r beta-mc-boxplot, fig.height = 5}
beta_cols <- c("#2E86C1", "#27AE60", "#F39C12", "#C0392B")
par(mfrow = c(1, 2))
boxplot(beta_wb ~ label, data = beta_mc_df,
  xlab = "Growth scenario", ylab = expression(hat(beta)),
  main = "Period 3 Weibull Shape by \u03b2 Scenario",
  col = beta_cols, outline = FALSE, las = 2, cex.axis = 0.75
)
boxplot(eta ~ label, data = beta_mc_df,
  xlab = "Growth scenario", ylab = expression(hat(eta)),
  main = "Period 3 Weibull Scale by \u03b2 Scenario",
  col = beta_cols, outline = FALSE, las = 2, cex.axis = 0.75
)
par(mfrow = c(1, 1))
```

```{r beta-mc-table}
beta_tbl <- do.call(rbind, lapply(beta_labels, function(lbl) {
  sub <- beta_mc_df[beta_mc_df$label == lbl, ]
  b   <- sub$beta_wb[!is.na(sub$beta_wb)]
  e   <- sub$eta[!is.na(sub$eta)]
  data.frame(
    Scenario = lbl,
    Valid    = length(b),
    Beta_med = round(median(b), 3),
    Beta_lo  = round(quantile(b, 0.025), 3),
    Beta_hi  = round(quantile(b, 0.975), 3),
    Eta_med  = round(median(e), 1),
    Eta_lo   = round(quantile(e, 0.025), 1),
    Eta_hi   = round(quantile(e, 0.975), 1)
  )
}))

knitr::kable(beta_tbl,
  caption = paste0(
    "Period 3 Weibull parameters by growth scenario (",
    n_sim_beta, " simulations each)"
  ),
  col.names = c(
    "Scenario", "Valid",
    "Med. \u03b2", "2.5% \u03b2", "97.5% \u03b2",
    "Med. \u03b7", "2.5% \u03b7", "97.5% \u03b7"
  ),
  row.names = FALSE
)
```

The median $\hat{\eta}$ rises monotonically as $\beta$ decreases from 1.4 to 0.3,
confirming that a stronger growth signal in the RGA model translates directly into
longer estimated characteristic lives in the downstream Weibull fit.  The $\beta = 1.1$
row sits just above the growth/degradation threshold: $\hat{\eta}$ is lower than
the two growth scenarios and reflects a slow leftward drift in the life distribution
across periods.

## Limitations

1. **Extrapolation range.** Reliability growth forecasts assume the improvement
   trend observed during testing continues into the future. The forecast
   window should be limited to a horizon where this assumption is defensible;
   large extrapolations beyond the test period carry substantial uncertainty.

2. **Fleet depletion.** Removing failed units from each successive period
   reduces the at-risk population. With small fleets or many predicted failures,
   the surviving pool for later periods may be too small to support meaningful
   Weibull estimation. Widening the intervals, increasing the initial fleet
   size, or modeling repairs/replacements explicitly can mitigate this.

3. **Homogeneous fleet.** `sim_failures()` uses a single runtime-proportional
   failure probability for all units. If the fleet contains distinct sub-populations
   (e.g., different software versions or duty cycles), separate simulations
   for each group should be run before pooling.

4. **Small sample Weibull fits.** Life distribution estimates are sensitive to
   sample size. With few failures, the confidence bounds on shape and scale
   will be wide; point estimates should be treated with caution and the full
   uncertainty interval reported.

## Conclusion

This analysis demonstrated a full reliability growth forecasting pipeline:
starting with a fitted Crow-AMSAA model, forecasting future failures, simulating
which units fail and when, and fitting Weibull distributions to the combined
historical and simulated failure data across multiple forecast periods. A Monte
Carlo analysis quantified the uncertainty in the Weibull parameters attributable
to the stochastic failure assignment process, while sensitivity analyses showed how 
failure count and growth model parameters affect estimation precision. This end-to-end 
workflow provides a practical framework for translating reliability growth signals
from testing into actionable fleet reliability insights.
