---
title: "ReliaGrowR: Open Source Software for Reliability Growth Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Keywords: Reliability Growth Analysis, ReliaGrowR, R package, reliability engineering, life data analysis

## Summary & Conclusions

Reliability Growth Analysis plays a crucial role in engineering and product development by improving system reliability throughout the development and testing phases. ReliaGrowR is an open-source software package developed specifically to support the analysis of reliability growth data. It offers a set of simple yet effective functions for analyzing failure data, estimating reliability parameters, and visualizing reliability trends over time. Built on the R programming language, widely used in statistical computing and data analysis, ReliaGrowR provides intuitive functionality suitable for both novice and experienced users. The package is lightweight and extensible, allowing for customization and enhancement, and making it a versatile tool for researchers and practitioners in the field of reliability engineering. 

## Introduction


## Implementation

ReliaGrowR is an R package designed for Reliability Growth Analysis (RGA), providing tools to analyze and visualize reliability growth data. The package includes functions for various reliability growth models, such as the Duane Model, Crow-AMSAA Model, Piecewise NHPP Model, and Piecewise NHPP with Change Point Detection. The package is built on the R programming language, which is widely used for statistical computing and data analysis.

The package is designed to be lightweight and easy to use, with a focus on providing essential functionality for RGA without unnecessary complexity. It is also designed to be extensible, allowing users to add custom models or features as needed. ReliaGrowR has one primary dependency on the `segmented` package [*] for regression modeling with break or change points, which is the underlying library for the Piecewise NHPP with or without change point detection.

## Usage

R is available on the Comprehensive R Archive Network (CRAN) [*]. To install R, follow the instructions provided on the CRAN website for the applicable operating system. Once R is installed, install the ReliaGrowR package from CRAN using the following command:

```{r install-package, eval=FALSE}
install.packages("ReliaGrowR")
```

To use the ReliaGrowR package, load the package into the current R session with the following command:

```{r load-package}
library(ReliaGrowR)
```

### The Duane Model

The Duane Model [*] provides a simple and graphical way to observe and analyze whether failure rates are improving as changes are made to a product or system. The Duane Model is a log-log plot of the cumulative Mean Time Between Failures (MTBF) vs cumulative time.

The slope of the line on the plot indicates the rate of reliability growth:

* A positive slope means that the system is improving (reliability is growing, the failure rate is decreasing).
* A zero slope means there is no change in reliability (the system is stable).
* A negative slope indicates that reliability is worsening (the failure rate is increasing).

To use the Duane Model in ReliaGrowR, use the `duane_plot` function. This function takes a a vector of failure times and a vector of failure counts, and generates a log-log plot of cumulative MTBF vs cumulative time.

First, set up some dummy cumulative time and failure data:

```{r echo=TRUE}
times <- c(100, 200, 300, 400, 500)
failures <- c(1, 2, 1, 3, 2)
```

Next, use the `duane_plot` function to create the plot:

```{r duane-plot, echo=TRUE}
fit <- duane_plot(times, failures)
```

The `duane_plot` function returns a `duane` object with the fitted model results that can be further customized or saved. The plot shows the cumulative MTBF on the y-axis and cumulative time on the x-axis, with a fitted line indicating the reliability growth trend.

### The Crow-AMSAA Model

The Army Materiel Systems Analysis Activity Model by Crow (Crow-AMSAA) [*] takes failure behavior as a Non-Homogeneous Poisson Process (NHPP) governed by a power law, making the model particularly effective for systems undergoing reliability growth due to continuous improvements.

Similar to the Duane Model, the shape of the model indicates the rate of reliability growth:

* A model fit with downward curvature means that the system is improving (reliability is growing, the failure rate is decreasing).
* A linear model means there is no change in reliability (the system is stable).
* A model fit with upward curvature indicates that reliability is worsening (the failure rate is increasing).

To use the Crow-AMSAA Model in ReliaGrowR, use the `rga` function. This function takes a vector of failure times and a vector of failure counts, and generates a plot of cumulative MTBF vs cumulative time with the fitted model.

First, set up some dummy cumulative time and failure data:

```{r echo=TRUE}
times <- c(100, 200, 300, 400, 500)
failures <- c(1, 2, 1, 3, 2)
```

Then use the `rga` function to fit the model and the `plot_rga` function to plot the results:

```{r}
result <- rga(times, failures)
plot_rga(result)
```

The `rga` function returns an `rga` object containing the fitted model parameters, and the `plot_rga` function generates a plot showing the cumulative MTBF on the y-axis and cumulative time on the x-axis, with a fitted curve indicating the reliability growth trend.

### The Piecewise NHPP Model

The Piecewise NHPP model [*], [*] is an extension of the standard NHPP model that includes different segments or phases of time that follow separate failure distributions. This model is particularly useful when a system experiences changes in failure behavior over different development phases, such as the initial, interim and final phases of a development process.

To use the Piecewise NHPP model in ReliaGrowR, first, set up some cumulative time and failure data and specify a breakpoint:

```{r echo=TRUE}
times <- c(25, 55, 97, 146, 201, 268, 341, 423, 513, 609, 710, 820, 940, 1072, 1217)
failures <- c(1, 1, 2, 4, 4, 1, 1, 2, 1, 4, 1, 1, 3, 3, 4)
breaks <- 500
```

Then use the `rga` function with model type "Piecewise NHPP model" to fit the model and the `plot_rga` function to plot the results:

```{r echo=TRUE}
result <- rga(times, failures, model_type = "Piecewise NHPP", breaks = breaks)
plot_rga(result)
```

### The Piecewise NHPP with Change Point Detection

The Piecewise NHPP with Change Point Detection [*] is an advanced model to identify changes in failure behavior and model system reliability. This method builds on the Piecewise NHPP model by introducing the concept of change points, which represent the time when the underlying failure behavior changes. Detection of change points involves statistical techniques that analyze failure data to automatically identify when the behavior changes, allowing for a more precise segmentation of the model into different distributions.

To use the Piecewise NHPP with Change Point Detection in ReliaGrowR, use the `rga` function with the model type set to "Piecewise NHPP" and breaks set to NULL. The function will automatically detect change points based on the provided failure data. First, set up some cumulative time and failure data:

```{r echo=TRUE}
times <- c(25, 55, 97, 146, 201, 268, 341, 423, 513, 609, 710, 820, 940, 1072, 1217)
failures <- c(1, 1, 2, 4, 4, 1, 1, 2, 1, 4, 1, 1, 3, 3, 4)
```

Then use the `rga` function with model type "Piecewise NHPP model" to fit the model and the `plot_rga` function to plot the results:

```{r echo=TRUE}
result <- rga(times, failures, model_type = "Piecewise NHPP")
plot_rga(result)
```

## Verification

ReliaGrowR was verified through unit tests and example analyses to ensure that the package performs as expected. The package includes a suite of tests that cover the core functionalities, including model fitting, plotting, and change point detection. These tests run automatically during package development to ensure reliability and correctness.

ReliaGrowR was also tested on different operating systems and R versions to ensure compatibility and performance. The results of these tests are documented on CRAN. Full documentation and working examples are available on the project website, where users can also contribute to or report issues with the package.

## Extensibility 

ReliaGrowR is designed to be extensible, allowing users to add custom models or features as needed.  The package has already been extended in a number of ways, including education [*], advanced visualization [*], and web-based applications [*]. Users can contribute to the package by submitting pull requests on the project repository, where the source code is hosted. The package is also open to contributions from the community, and users are encouraged to report issues or suggest improvements.

## References


