---
title: "A Simulation-based Framework for Growth-Adjusted Life Distribution Forecasting"
author: "Paul Govan, GE Aerospace"
date: "Key Words: Reliability Growth Analysis, Life Forecasting, Monte Carlo
  Simulation, Weibull Analysis"
output:
  word_document:
    reference_docx: "rams-reference.docx"
    fig_caption: true
    fig_width: 3.2
    fig_height: 2.6
  html_document:
    df_print: paged
bibliography: references.bib
csl: ieee.csl
link-citations: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
library(ReliaGrowR)
library(WeibullR)
```


## Summary & Conclusions

Reliability growth analysis (RGA) quantifies how system reliability improves as design
changes are introduced during development, but translating the observed growth trend into
a fleet-level life distribution normally requires several disconnected analyses. This article presents an integrated workflow that connects RGA to life estimation through stochastic simulation, implemented as a reproducible, open-source workflow.
The example is structured in two phases. A
developmental growth test of sequential prototypes, in which design improvements
lead to an improvement trend.
This trend is then projected onto a fielded fleet of the matured design: over a fixed
forecast window the growth model lowers the number of additional failures expected from
the fleet's own life distribution; a conditional Weibull simulation generates failure
times for the surviving units over that window; and the simulated failures are combined
with the observed fleet data to fit a growth-adjusted Weibull life distribution. A parallel Weibull fitted to the
fleet data alone provides a no-growth baseline for comparison. Monte Carlo analysis
quantifies the variability introduced by growth-parameter uncertainty and the simulation,
and sensitivity analyses examine the dependence on the fleet assessment point and
the strength of the growth trend. By construction, incorporating the growth trend shifts
the characteristic life upward relative to the no-growth baseline. The contribution of the
workflow is to formalize and quantify the growth-induced shift in characteristic life by
propagating both growth-parameter uncertainty and simulation variability into the forecast;
establishing whether this adjustment improves predictive accuracy over no-growth baselines
requires validation against a fleet exhibiting true continued growth, which is left to future
work. This enables decision-makers to assess whether the observed developmental trend
justifies an upward revision of fleet-level reliability estimates and to quantify the
forecast's sensitivity to the assumed growth continuation. The method provides a repeatable
means of propagating observed reliability growth into fleet-level life distribution estimates.
Because the projection assumes the developmental growth trend continues to be reflected in
field reliability, it is designed to address the potential underestimation of future
reliability that may result when growth trends are not propagated into fleet-level forecasts.
The forecast should be read as conditional on the developmental growth trend continuing in
service. When this assumption is justified, such as when design improvements are expected to
persist into fielding, the method provides a quantitative, reproducible means of propagating
that growth into fleet-level life estimates, with explicit variability bounds that fragmented
analyses do not deliver.

## Introduction

Reliability growth analysis quantifies how system reliability improves during
development, yet translating the observed growth trend into a fleet-level life
distribution is rarely formalized. The challenge is compounded by limited life data,
particularly for high-reliability, long-life products for which
collecting sufficient failure information is costly and time-consuming. As product
quality improves, the likelihood of
observing failures within a practical test window decreases, creating a data-scarcity
problem that undermines traditional reliability estimation. The difficulty is
compounded by censoring: units are routinely removed from test before failing, or
their failure times are known only to fall within an interval, so the effective
information content of a campaign is smaller than the sample size suggests
[@zhang2013censored; @alqasem2025phn].

The consequences of analyzing limited, heavily censored data without appropriate
methods are well documented. Maximum likelihood estimation becomes unreliable with
small samples, producing biased and highly variable parameter estimates, and the
shape parameter of the Weibull distribution is especially sensitive
[@makalic2023mml]. Most critically, reliability
predictions that extrapolate beyond the observed data become unreliable, with
serious safety and economic implications [@lou2023acceleration]. A range of
remedies has been proposed, from integrating historical degradation data
[@kang2023wiener] to generalized inference for remaining useful life
[@yan2025rul] and to case-specific Weibull fits for sparse field data
[@grile2024satellite].

A distinct but under-addressed problem arises when the system is actively being improved
during test: the life distribution itself is not fixed. Existing remedies for data scarcity
improve the estimation of a static distribution but do not propagate the observed reliability
growth trend into fleet-level life predictions. RGA characterizes the system-level failure
intensity and is conventionally used to project a future failure count or MTBF. However,
this system-level trend is typically analyzed in isolation from unit-level life estimation.
As a result, the projected improvement is not reflected in the fleet-level life distribution,
and reliability forecasts may systematically underestimate future performance.

This article closes this gap by integrating RGA with conditional Weibull simulation. In
contrast to applying
a fixed reliability-growth multiplier to a baseline distribution, the growth model here
adjusts the failure count expected over a fixed forecast window, a conditional simulation
propagates the growth-adjusted failure behavior onto the surviving fleet, and a
growth-adjusted Weibull distribution is
fitted to the combined data and compared against a no-growth baseline, with the
variability of the projection quantified by Monte Carlo. This integrated approach is intended
to propagate both growth-parameter uncertainty and simulation variability more transparently
than ad hoc multipliers, though empirical comparison of predictive accuracy is left to future
work. The forecast is explicitly
conditional on the observed growth continuing, where fragmented analyses risk
underestimating future reliability, an over-extrapolated growth trend can equally
overestimate it, and the framework is intended to make that dependence visible rather
than to assert a single point forecast. The complete workflow is implemented in
open-source [@govan2026reliagrowr; @govan2024reliagrowrpkg], with Weibull fitting
performed by maximum likelihood estimation (MLE) [@silkworth2025weibullr].

## Simulation Framework

The framework integrates three sequential stages: growth-trend estimation,
conditional stochastic simulation, and growth-adjusted life estimation, into a
unified pipeline that propagates an observed reliability growth trend into a
fleet-level Weibull life distribution. The inputs come from two phases of a program, both involving non-repairable
units. A *developmental growth test* records the cumulative test times of sequential
prototypes built under a test-analyze-fix process, and a
Crow-AMSAA model (Eqs. 8--11) quantifies the system-level improvement
trend. A *fielded fleet* of the resulting matured design then provides failure and
suspension times over a finite campaign of duration $T_{\text{end}}$, to which a
two-parameter Weibull distribution (Eq. 7) is fitted to characterize the unit-level
failure behavior and to serve as the no-growth baseline. Together, the fitted growth
model and the fitted Weibull supply the inputs needed for the two subsequent stages.

**Stage 1 --- Forecast window and growth-adjusted failure count.** The forecast is run
over a fixed *forecast window* $w$, a forward increment applied to each
surviving unit. Over this window the
fitted fleet Weibull (shape $\beta$, scale $\eta$) is expected, with no growth, to
produce

$$F_{0} = \sum_{i=1}^{n_{s}} \left[1 - \exp\!\left\{-\left[
\left(\tfrac{r_{i}+w}{\eta}\right)^{\beta} -
\left(\tfrac{r_{i}}{\eta}\right)^{\beta}\right]\right\}\right] \tag{1}$$

failures among the surviving units, where $r_{i}$ is each unit's current age.

Reliability growth lowers this count: an improving fleet accumulates failures more slowly.
Re-anchoring the Crow-AMSAA scale to the fleet ($\lambda = N_{\text{current}}/T_{\text{end}}^{\,\beta_{g}}$),
the expected additional failures over the cumulative window $\Delta t = n_{s}\,w$ are
$N_{\text{current}}[(1+\rho)^{\beta_{g}} - 1]$ with $\rho = \Delta t / T_{\text{end}}$. The
same since-inception anchoring enters the no-growth ($\beta_{g}=1$) case, so its bias
cancels when the growth effect is expressed as the *ratio* of the two,

$$c(\beta_{g}) = \frac{(1+\rho)^{\beta_{g}} - 1}{(1+\rho) - 1}
= \frac{(1+\rho)^{\beta_{g}} - 1}{\rho}, \tag{2}$$

which satisfies $c(1)=1$ and decreases as the growth strengthens
($\beta_{g}\!\downarrow$). The growth-adjusted target failure count over the window is then

$$n = F_{0}\,c(\beta_{g}). \tag{3}$$

Because $c(1)=1$, with no growth the target reduces to the forward-correct benchmark
$F_{0}$ and the calibrated scale (Stage 2) returns the fitted fleet value; the growth
effect enters only through $c(\beta_{g}) < 1$, which forces a larger scale to reproduce
the reduced count. The projection thus assumes only that the *relative* growth reduction
estimated in development carries over to the fleet. Because the power law is scale-free, only
the fractional increment $\rho$ and the dimensionless $\beta_{g}$ enter $c$, so no
developmental time unit is imported into the fleet forecast.

**Stage 2 --- Conditional Weibull simulation.** Each surviving unit enters the
simulation window with a known current runtime $r_{i}$. The probability of failing
within the window $[r_i, r_i + w]$, conditional on having survived to $r_i$, is

$$p_{i} = 1 - \exp\!\left\{-\left[\left(\tfrac{r_{i}+w}{\eta}\right)^{\beta} -
\left(\tfrac{r_{i}}{\eta}\right)^{\beta}\right]\right\}. \tag{4}$$

The scale parameter $\eta$ is calibrated so that the expected number of conditional
failures over the window matches the target count $n$:

$$\sum_{i=1}^{n_{s}} p_{i}(\eta) = n, \tag{5}$$

which is solved numerically for $\eta$. The calibrated $\eta$ reflects reliability
growth: because growth lowers the target count $n$ below the no-growth benchmark $F_{0}$,
a larger characteristic life is required to reproduce the reduced count over the same
window. This upward shift in $\eta$ is therefore a *mechanistic* consequence of the
growth-reduced failure count rather than an independently discovered effect; the
framework's role is to quantify its magnitude and variability, not to establish its
existence. Units are then sampled with
probability proportional to $p_{i}$, and a
failure time for each selected unit is drawn by inverse-transform sampling of the
truncated conditional Weibull,

$$t_{i} = \eta\left[\left(\tfrac{r_{i}}{\eta}\right)^{\beta} -
\ln\!\Big(1 - U_{i}\big(1 - e^{-\Delta h_{i}}\big)\Big)\right]^{1/\beta}, \tag{6}$$

where $U_{i}\sim\text{Uniform}(0,1)$ and
$\Delta h_{i} = \big((r_{i}+w)/\eta\big)^{\beta} - (r_{i}/\eta)^{\beta}$. The
non-selected units are right-censored at $r_{i}+w$. The procedure is summarized in
Algorithm 1.

> **Algorithm 1 --- Conditional Weibull failure simulation**
> 1. Calibrate the scale $\eta$ by solving Eq. (5) for the target count $n$.
> 2. Compute the per-unit window failure probability $p_{i}$, Eq. (4).
> 3. Sample $n$ units with probability proportional to $p_{i}$.
> 4. Draw a failure time for each selected unit by inverse transform, Eq. (6).
> 5. Label the remaining units as suspensions at $r_{i}+w$.

**Stage 3 --- Growth-adjusted Weibull fit and Monte Carlo variability assessment.**
The simulated failure and suspension times are combined with the observed fleet failures
to form an augmented dataset. A Weibull is fitted to obtain the growth-adjusted life
distribution, and a parallel Weibull fitted to the fleet data alone serves as the
no-growth baseline. The growth effect manifests primarily as an increase in the estimated
characteristic life $\eta$: the simulated failures occur later than they would under the
no-growth failure rate, shifting the distribution rightward. The headline comparison fits a two-parameter Weibull by maximum likelihood
to both the baseline and the growth-augmented data, so the two differ only in their data, and
reports the B10 life as the summary indicator. To characterize the variability in the
forecast, Stages 1 and 2 are repeated $M$ times: each iteration resamples the estimated
growth parameter
$\beta_{g}$ from its sampling distribution and recomputes the growth-adjusted count $n$
(Eqs. 2--3), redraws the conditional failures over the window, and refits the Weibull, so that
both growth-parameter estimation uncertainty and simulation randomness are reflected in
the resulting empirical distributions of $\hat{\beta}$ and $\hat{\eta}$, which are
summarized by medians and 95% percentile intervals.

## Case Study

The case study proceeds in two steps. First, a reliability growth analysis fits the
Crow-AMSAA model to a *developmental growth test* in which prototypes are built and tested
to failure in sequence. Second, a Weibull life analysis characterizes a *fielded fleet* of
100 units of the matured design, each observed over a 1000 time-unit campaign. The growth
trend estimated in the first step is then projected onto the fleet to forecast its life
distribution. Both phases involve non-repairable units.

The failure mechanism is wear-out, so the hazard rate
increases with age. To provide a reproducible example, the failure and right-censored
suspension times are simulated from the matured Weibull population (shape 4, scale 1456,
random seed 8): each unit is assigned a lifetime, units that fail before the end of the
campaign are recorded as exact failures, and the remainder are right-censored at the end
of the campaign.

In this illustrative example, the no-growth baseline fitted
below coincides with the ground-truth life distribution, and the growth-adjusted forecast
developed in the later sections is the *optimistic bound* that would apply only if the
developmental improvement trend continued into service. The example is intended to
demonstrate the mechanics and magnitude of that growth adjustment, not its predictive
accuracy against a fleet that truly keeps improving (see the Limitations section).

```{r data}
set.seed(8)
n_units <- 100
test_end <- 1000
beta_mech <- 4 # wear-out shape of the matured design
eta_matured <- 1456 # characteristic life of the matured design
fleet_life <- eta_matured * (-log(runif(n_units)))^(1 / beta_mech)

failures <- sort(fleet_life[fleet_life <= test_end])
suspensions <- rep(test_end, sum(fleet_life > test_end))

n_test_failures <- length(failures)
n_surviving <- length(suspensions)
```

This draw yields `r n_test_failures` exact failures, leaving `r n_surviving` surviving
units at risk during the forecast period.

A two-parameter Weibull distribution is fitted to the failure and suspension data
using MLE, providing the no-growth baseline life
distribution. The Weibull cumulative distribution function is

$$F(t) = 1 - \exp\!\left[-\left(\tfrac{t}{\eta}\right)^{\beta}\right], \qquad t > 0, \tag{7}$$

where $\beta$ is the shape parameter and $\eta$ is the scale (characteristic life).

```{r weibull-fit, fig.cap = "Weibull fit to the fielded fleet (no-growth baseline)."}
obj <- wblr(failures, suspensions,
  col = "black", label = "Fielded Fleet", is.plot.cb = FALSE
)
obj <- wblr.fit(obj, method.fit = "mle")
sim_beta <- obj$fit[[1]]$beta
eta_orig <- obj$fit[[1]]$eta
plot(obj,
  main = "Weibull Fit to Fielded Fleet (No-Growth Baseline)",
  xlab = "Operating Time", ylab = "Unreliability"
)
```

Figure 1 shows the Weibull fit to the fielded fleet with no growth. The estimated shape
parameter is $\beta$ = `r round(sim_beta, 3)` and the scale
parameter is $\eta$ = `r round(eta_orig, 1)`. A shape parameter greater than one
confirms the wear-out failure mode. This Weibull $\beta$ describes the unit-level hazard
in *individual age* and is distinct from the Crow-AMSAA growth parameter $\beta_g$
introduced next, which describes the system-level failure-intensity trend.

## Reliability Growth Analysis

```{r rga-convert}
set.seed(3)
N_dev <- 30 # prototypes built and tested in sequence
g_exp <- 0.8 # strength of design-life improvement with build index

eta_build <- eta_matured * (seq_len(N_dev) / N_dev)^g_exp
dev_life <- eta_build * (-log(runif(N_dev)))^(1 / beta_mech)
dev_cum <- cumsum(dev_life)
```

A Crow-AMSAA non-homogeneous Poisson process (NHPP) model [@crow1975reliability]
is fitted to the developmental record. The expected
cumulative number of failures by cumulative operating time $t$ is

$$N(t) = \lambda\, t^{\beta_{g}}, \tag{8}$$

with corresponding failure intensity

$$u(t) = \frac{dN(t)}{dt} = \lambda\,\beta_{g}\,t^{\beta_{g}-1}. \tag{9}$$

By default the parameters are estimated by ordinary least squares on the linearized
form of (8),

$$\ln N(t) = \ln \lambda + \beta_{g}\ln t. \tag{10}$$

Alternatively, the parameters may be estimated by MLE. The reliability
growth rate is

$$g = 1 - \beta_{g}. \tag{11}$$

A value $\beta_{g} < 1$ (equivalently $g > 0$) indicates reliability improvement.

```{r rga-fit}
dev_fit <- rga(
  data.frame(times = dev_cum, failures = rep(1, N_dev)),
  times_type = "cumulative_failure_times"
)
growth_beta <- as.numeric(dev_fit$betas)
dev_r2 <- summary(dev_fit$model)$r.squared
```

```{r rga-plot, results = "asis", fig.cap = "Crow-AMSAA reliability growth model fitted to the developmental test."}
plot(dev_fit,
  main = "Developmental Reliability Growth Test",
  xlab = "Cumulative Test Time", ylab = "Cumulative Failures"
)
```

Figure 2 shows the reliability growth model for the developmental phase. The fitted
growth parameter is $\beta_{g}$ = `r round(growth_beta, 3)`
($R^2$ = `r round(dev_r2, 3)`), indicating a decreasing failure intensity as the
design matured, and the reliability growth
rate (11) is $g = 1 - \beta_{g}$ = `r round(1 - growth_beta, 3)`.

## Reliability Growth Forecast

The growth trend is projected onto the fielded fleet following the Stage 1 procedure of
the Simulation Framework section. The forecast is run over a fixed *forecast window*. Over that window the fitted fleet Weibull sets the no-growth failure
count, and the growth *rate* $\beta_{g}$ imported from the developmental test lowers it
through the Crow-AMSAA count-ratio. Re-anchoring the Crow-AMSAA scale to the fleet
($\lambda = N_{\text{current}}/T_{\text{end}}^{\,\beta_{g}}$, with $T_{\text{end}}$ the
fleet's cumulative operating time and $N_{\text{current}}$ the failures observed so far)
makes that ratio pass through the fleet's observed state. The forecast is therefore
conditional on the developmental growth trend continuing to be reflected in field
reliability.

```{r forecast-calc}
# Forecast window: a modest forward increment (~40% of the observation period) applied to
# each surviving unit, chosen rather than solved for.
forecast_window <- 400
test_end_cum_time <- sum(c(failures, suspensions))

# No-growth count the fitted fleet Weibull expects over the window (forward hazard).
expected_failures <- function(w, eta, r, beta) {
  sum(1 - exp(-(((r + w) / eta)^beta - (r / eta)^beta)))
}
# Scale that reproduces a target expected count over the window (inverts Eq. 5).
solve_eta <- function(target, r, beta, window) {
  uniroot(function(eta) expected_failures(window, eta, r, beta) - target,
    interval = c(1e-3, 1e9)
  )$root
}
# Crow-AMSAA growth count-ratio: growth lowers the expected count; equals 1 at beta_g = 1.
count_ratio <- function(bg, rho) ((1 + rho)^bg - 1) / rho

f0 <- expected_failures(forecast_window, eta_orig, rep(test_end, n_surviving), sim_beta)
rho <- (n_surviving * forecast_window) / test_end_cum_time
n_expected <- f0 * count_ratio(growth_beta, rho)
n_forecast <- round(n_expected)
n_target <- n_test_failures + n_forecast
```

Figure 3 shows the reliability growth forecast beyond the developmental phase. Over the
forecast window of $w$ = `r forecast_window` operating time units, the fitted fleet Weibull
expects $F_{0}$ = `r round(f0, 1)` failures with no growth (Eq. 1). The Crow-AMSAA growth
count-ratio (Eq. 2) is $c$ = `r round(count_ratio(growth_beta, rho), 2)`, lowering the
growth-adjusted target to $n$ = `r n_forecast` additional failures (Eq. 3), for a total of
`r n_target` cumulative failures.

```{r forecast-plot, warning = FALSE, fig.cap = "Developmental Crow-AMSAA growth model fitted to the sequential prototype builds and forecast forward along the same trend, with confidence bounds."}
dev_target <- N_dev + n_forecast
t_f_dev <- (dev_target / as.numeric(dev_fit$lambdas))^(1 / growth_beta)
dev_forecast_times <- seq(max(dev_cum) * 1.001, t_f_dev, length.out = 50)

fc <- predict_rga(dev_fit, times = dev_forecast_times)
plot(fc,
  main = "Reliability Growth Forecast",
  xlab = "Cumulative Test Time", ylab = "Cumulative Failures"
)
```

## Simulating Failures

Conditional Weibull failure times for the `r n_surviving` surviving units are
generated using Algorithm 1. Each unit
has already accumulated 1000 operating time units; failure times are therefore drawn
from the Weibull distribution conditional on the current age over the forecast window
$w$.

```{r simulate}
set.seed(123)
# Calibrate the scale to the growth-adjusted expected count over the window, then draw the
# integer number of failures; at beta_g = 1 this returns the fitted fleet scale exactly.
sim_eta <- solve_eta(n_expected, rep(test_end, n_surviving), sim_beta, forecast_window)
sim_result <- sim_failures(
  n = n_forecast,
  runtimes = rep(1000, n_surviving),
  window = forecast_window,
  beta = sim_beta,
  eta = sim_eta
)
```

The calibrated scale parameter is $\eta$ = `r round(sim_eta, 1)`, compared with the
no-growth fleet estimate of $\eta$ = `r round(eta_orig, 1)`. The increase is the
mechanical consequence of the assumed growth: growth lowers the expected failure count
over the fixed window, so a larger characteristic life is required to reproduce the
reduced count. Of the
`r n_surviving` surviving units, `r n_forecast` receive simulated failure times and
`r n_surviving - n_forecast` are right-censored at the end of the window. The
simulated failures are combined with the `r n_test_failures` observed fleet failures.

```{r combine}
sim_fail_times <- sim_result$runtime[sim_result$type == "Failure"]
sim_susp_times <- sim_result$runtime[sim_result$type == "Suspension"]
combined_failures <- c(failures, sim_fail_times)
combined_suspensions <- sim_susp_times
```

### Weibull Comparison

Two Weibull models are fitted to contrast the effect of reliability growth. The
*without-growth* model is fitted to the fielded fleet data alone
(`r n_test_failures` failures, `r n_surviving` suspensions at $t = 1000$),
representing the life distribution implied by the fleet without projecting the growth
trend forward. The *with-growth* model is fitted to the combined dataset
(`r n_test_failures` fleet failures, `r n_forecast` simulated failures, and
`r n_surviving - n_forecast` suspensions), representing the life distribution when
growth-adjusted failure times are included. Table 1 shows the parameters and B10 life for
the two Weibull models, and Figure 4 shows the Weibull plot comparison.

```{r comparison-plot, fig.cap = "Weibull comparison with and without reliability growth."}
# Headline comparison is like-for-like: a 2-parameter Weibull fitted by MLE to both the
# fielded fleet (no-growth baseline) and the growth-augmented data, so the two fits differ
# only in their data, not in parameterization or estimation method.
obj_nogrowth <- wblr(failures, suspensions,
  col = "grey50", pch = 1, lty = 2,
  label = "Without Growth", is.plot.cb = FALSE
)
obj_nogrowth <- wblr.fit(obj_nogrowth, method.fit = "mle")

obj_growth <- wblr(combined_failures, combined_suspensions,
  col = "black", pch = 16, lty = 1,
  label = "With Growth (2P)", is.plot.cb = FALSE
)
obj_growth <- wblr.fit(obj_growth, method.fit = "mle")

plot.wblr(list(obj_nogrowth, obj_growth),
  main = "Weibull Comparison: With vs. Without Reliability Growth",
  is.plot.legend = TRUE
)
```

```{r comparison-table}
growth_wb_beta <- obj_growth$fit[[1]]$beta
growth_wb_eta <- obj_growth$fit[[1]]$eta
nogrowth_wb_beta <- obj_nogrowth$fit[[1]]$beta
nogrowth_wb_eta <- obj_nogrowth$fit[[1]]$eta

# B10 life (10th-percentile life): t such that F(t) = 0.10.
b10 <- function(beta, eta, t0 = 0) t0 + eta * (-log(0.90))^(1 / beta)
nogrowth_b10 <- b10(nogrowth_wb_beta, nogrowth_wb_eta)
growth_b10 <- b10(growth_wb_beta, growth_wb_eta)

params <- data.frame(
  Scenario = c("Without Growth (2P)", "With Growth (2P)"),
  Beta = round(c(nogrowth_wb_beta, growth_wb_beta), 3),
  Eta = round(c(nogrowth_wb_eta, growth_wb_eta), 1),
  B10 = round(c(nogrowth_b10, growth_b10), 1)
)

knitr::kable(params,
  caption = "Weibull parameters and B10 life without and with reliability growth (two-parameter MLE)",
  col.names = c("Scenario", "β", "η", "B10"),
  row.names = FALSE
)
```

## Monte Carlo Analysis

A single simulation produces one realization of the `r n_forecast` forecast failure
times; different random draws yield different combined datasets and therefore different
Weibull estimates. To characterize this variability, the simulation is repeated
`r 500` times following the Monte Carlo procedure of the Simulation Framework section.
Two sources of variability are propagated. Each iteration draws $\beta_g$ from its
sampling distribution and recomputes the growth-adjusted failure count over the fixed
window from Eqs. (2)--(3); it then redraws the forecast failures, combines them with the
fleet failures, and refits a Weibull distribution with both parameters freely estimated. In
this example the developmental trend is estimated precisely, the cumulative-data fit
pins $\beta_g$ down with a small standard error, so the resulting spread is dominated
by the conditional simulation, with growth-parameter uncertainty included for
completeness but contributing little. The Monte Carlo and
sensitivity loops fit two-parameter Weibulls by maximum likelihood, matching the headline
comparison, so the B10 life is directly comparable throughout.

```{r mc-helpers}
# Given a target expected count over the window, calibrate the scale, draw round(target)
# failures, combine with the observed failures, and refit a two-parameter Weibull.
sim_and_fit <- function(test_failures, target_exp, runtimes, window, beta) {
  eta_c <- solve_eta(target_exp, runtimes, beta, window)
  sim_i <- sim_failures(
    n = round(target_exp), runtimes = runtimes, window = window,
    beta = beta, eta = eta_c
  )
  sim_f <- sim_i$runtime[sim_i$type == "Failure"]
  sim_s <- sim_i$runtime[sim_i$type == "Suspension"]
  obj_i <- wblr(c(test_failures, sim_f), sim_s, is.plot.cb = FALSE)
  obj_i <- wblr.fit(obj_i, method.fit = "mle")
  data.frame(beta = obj_i$fit[[1]]$beta, eta = obj_i$fit[[1]]$eta)
}
```

```{r mc-loop, results = "hide"}
set.seed(99)
n_mc <- 500
mc_results <- vector("list", n_mc)

# Sampling distribution of the developmental growth parameter.
betag_hat <- as.numeric(dev_fit$betas)
betag_se <- as.numeric(dev_fit$betas_se)

for (i in seq_len(n_mc)) {
  tryCatch(
    {
      # Draw a growth parameter and recompute the growth-adjusted count over the window.
      beta_i <- rnorm(1, betag_hat, betag_se)
      if (beta_i <= 0) stop("non-positive growth parameter draw")
      target_i <- f0 * count_ratio(beta_i, rho)
      if (!is.finite(target_i) || target_i <= 0) stop("non-positive count draw")

      mc_results[[i]] <- sim_and_fit(
        failures, target_i, rep(1000, n_surviving), forecast_window, sim_beta
      )
    },
    error = function(e) NULL
  )
}

mc_df <- do.call(rbind, Filter(Negate(is.null), mc_results))
```

The fitted $\hat{\beta}$, $\hat{\eta}$, and B10 life across all valid iterations are
summarized in Table 2; the median, mean, and 95% percentile interval describe
the spread introduced by the propagated growth-parameter uncertainty and the stochastic
simulation, with the no-growth baseline shown for reference.

```{r mc-table}
mc_df$b10 <- b10(mc_df$beta, mc_df$eta)

mc_summary <- data.frame(
  Parameter = c("β", "η", "B10"),
  NoGrowth = round(c(nogrowth_wb_beta, nogrowth_wb_eta, nogrowth_b10), 3),
  Median = round(c(median(mc_df$beta), median(mc_df$eta), median(mc_df$b10)), 3),
  Mean = round(c(mean(mc_df$beta), mean(mc_df$eta), mean(mc_df$b10)), 3),
  CI_lo = round(c(
    quantile(mc_df$beta, 0.025),
    quantile(mc_df$eta, 0.025),
    quantile(mc_df$b10, 0.025)
  ), 3),
  CI_hi = round(c(
    quantile(mc_df$beta, 0.975),
    quantile(mc_df$eta, 0.975),
    quantile(mc_df$b10, 0.975)
  ), 3)
)

knitr::kable(mc_summary,
  caption = paste0(
    "Monte Carlo summary of Weibull parameters and B10 life (",
    nrow(mc_df), " valid iterations)"
  ),
  col.names = c(
    "Parameter", "No Growth", "Median", "Mean", "2.5%", "97.5%"
  ),
  row.names = FALSE
)
```

The Monte Carlo distributions quantify how much the growth-adjusted Weibull estimates
vary across repeated simulations, here almost entirely from the stochastic simulation
rather than from growth-parameter estimation uncertainty. These bands represent the
variability of the forecast conditional on the fitted growth model and the assumed growth
continuation, not inferential confidence intervals. They quantify the repeatability of the
method and the sensitivity of the forecast to simulation randomness, but they do not assess
the accuracy of the growth assumption itself.
They should be read as forecast variability bands conditional on the fitted model and
the assumed growth scenario, not as inferential confidence intervals or a hypothesis
test: in particular, model-form uncertainty (the choice of the Crow-AMSAA power law) is
not represented. The separation between the no-growth baseline and the Monte Carlo
distribution of $\hat{\eta}$ reflects the genuine growth reduction in the failure count
($c(\beta_{g}) < 1$); with no growth the target reduces to $F_{0}$ and the distribution
centers on the baseline.

## Sensitivity Analysis

The Monte Carlo analysis held the fleet assessment and the input shape parameter
fixed while varying the simulated draws. This section examines how two key inputs
affect the growth-adjusted Weibull: the fleet assessment point, and the
strength of the growth trend. Each scenario runs `r 200` iterations.

### Effect of Fleet Assessment Point

The point at which the fleet is assessed sets how many failures have been observed and how
the baseline fleet Weibull is fitted, which feed the no-growth count $F_{0}$ over the
window. The same fleet is assessed at four service points, yielding progressively more
observed failures.

```{r sens-fail-setup}
assessment_times <- c(850, 950, 1050, 1150)
fleet_obs <- vapply(assessment_times, function(te) sum(fleet_life <= te), integer(1))
fleet_labels <- paste0(fleet_obs, " failures")
```

```{r sens-fail-mc, results = "hide"}
set.seed(42)
n_mc_sens <- 200

sens_fail_list <- lapply(seq_along(assessment_times), function(k) {
  te <- assessment_times[k]
  fail_te <- sort(fleet_life[fleet_life <= te])
  susp_te <- rep(te, sum(fleet_life > te))
  n_obs <- length(fail_te)
  n_surv <- length(susp_te)

  obj_te <- wblr.fit(wblr(fail_te, susp_te, is.plot.cb = FALSE), method.fit = "mle")
  beta_te <- obj_te$fit[[1]]$beta
  eta_te <- obj_te$fit[[1]]$eta

  if (n_surv <= 0) {
    return(NULL)
  }
  # No-growth count over the fixed window from this assessment's own fitted baseline,
  # lowered by the growth count-ratio at this assessment's cumulative exposure.
  f0_te <- expected_failures(forecast_window, eta_te, rep(te, n_surv), beta_te)
  rho_te <- (n_surv * forecast_window) / sum(c(fail_te, susp_te))
  target_te <- f0_te * count_ratio(growth_beta, rho_te)
  if (!is.finite(target_te) || target_te <= 0) {
    return(NULL)
  }

  rows <- lapply(seq_len(n_mc_sens), function(i) {
    tryCatch(
      {
        r <- sim_and_fit(fail_te, target_te, rep(te, n_surv), forecast_window, beta_te)
        cbind(scenario = fleet_labels[k], r)
      },
      error = function(e) NULL
    )
  })
  do.call(rbind, Filter(Negate(is.null), rows))
})

sens_fail_df <- do.call(rbind, Filter(Negate(is.null), sens_fail_list))
sens_fail_df$scenario <- factor(sens_fail_df$scenario, levels = fleet_labels)
```

```{r sens-fail-boxplot, fig.height = 3.6, fig.width = 5.0, fig.cap = "Growth-adjusted Weibull parameters by fleet assessment point."}
par(mfrow = c(1, 2), mar = c(7, 4, 3, 1))

fail_cols <- c("white", "grey80", "grey55", "grey30")
boxplot(beta ~ scenario,
  data = sens_fail_df,
  col = fail_cols, outline = FALSE,
  main = "Fitted β by Failures Observed",
  xlab = "", ylab = expression(hat(beta)),
  las = 2, cex.axis = 0.8
)

boxplot(eta ~ scenario,
  data = sens_fail_df,
  col = fail_cols, outline = FALSE,
  main = "Fitted η by Failures Observed",
  xlab = "", ylab = expression(hat(eta)),
  las = 2, cex.axis = 0.8
)

par(mfrow = c(1, 1))
```

Figure 5 shows the Weibull parameters by increasing failures observed. Across the four
assessment points the growth count-ratio is essentially constant ($c \approx 0.8$), so the
growth adjustment itself is robust to when the fleet is assessed; the visible variation is
inherited from the *baseline* Weibull fit, which is depressed by the heavy censoring of an
early assessment (`r fleet_obs[1]` observed failures) and stabilizes once enough failures have accrued.
The practical implication is that a dependable growth forecast depends less on the
assessment timing than on having enough observed failures to pin down the fleet's baseline
life distribution.

### Effect of Growth Parameter

The growth parameter $\beta_{g}$ controls the growth count-ratio $c(\beta_{g})$:
stronger growth (smaller $\beta_{g}$) lowers the expected failure count over the fixed
window, producing fewer simulated failures and a larger fitted $\eta$. Four growth
scenarios are examined while holding all other inputs fixed; each scales the same
no-growth count $F_{0}$ by its ratio $c(\beta_{g})$, so the scenarios differ only in the
growth strength.

```{r growth-setup}
growth_scenarios <- c(0.4, 0.6, 0.8, 1.0)
growth_labels <- c(
  "βg = 0.4 (strong)",
  "βg = 0.6 (moderate)",
  "βg = 0.8 (mild)",
  "βg = 1.0 (none)"
)
```

```{r growth-mc, results = "hide"}
set.seed(77)

sens_growth_list <- lapply(seq_along(growth_scenarios), function(k) {
  gb <- growth_scenarios[k]

  # Lower the fixed-window no-growth count by the growth ratio for this scenario.
  target_k <- f0 * count_ratio(gb, rho)

  if (target_k <= 0) {
    return(NULL)
  }

  rows <- lapply(seq_len(n_mc_sens), function(i) {
    tryCatch(
      {
        r <- sim_and_fit(
          failures, target_k, rep(1000, n_surviving), forecast_window, sim_beta
        )
        cbind(scenario = growth_labels[k], r)
      },
      error = function(e) NULL
    )
  })
  do.call(rbind, Filter(Negate(is.null), rows))
})

sens_growth_df <- do.call(rbind, Filter(Negate(is.null), sens_growth_list))
sens_growth_df$scenario <- factor(sens_growth_df$scenario, levels = growth_labels)
```

```{r growth-boxplot, fig.height = 3.6, fig.width = 5.0, fig.cap = "Growth-adjusted Weibull parameters by growth strength."}
par(mfrow = c(1, 2), mar = c(8, 4, 3, 1))

growth_cols <- c("white", "grey80", "grey55", "grey30")
boxplot(beta ~ scenario,
  data = sens_growth_df,
  col = growth_cols, outline = FALSE,
  main = "Fitted β by Growth Strength",
  xlab = "", ylab = expression(hat(beta)),
  las = 2, cex.axis = 0.65
)

boxplot(eta ~ scenario,
  data = sens_growth_df,
  col = growth_cols, outline = FALSE,
  main = "Fitted η by Growth Strength",
  xlab = "", ylab = expression(hat(eta)),
  las = 2, cex.axis = 0.65
)

par(mfrow = c(1, 1))
```

Figure 6 shows the Weibull parameters by increasing growth parameter. Stronger reliability
growth (smaller $\beta_{g}$) lowers the expected failure count over the fixed window
through the count-ratio $c(\beta_{g})$, producing fewer simulated failures and a larger
fitted $\eta$. At $\beta_{g} = 1$ (no growth) the ratio $c(1) = 1$ leaves the target at
the forward-correct no-growth count $F_{0}$, and the growth-adjusted distribution recovers
the no-growth baseline --- its median $\eta$ returns to the fitted fleet value up to Monte
Carlo noise. The growth-adjusted $\eta$ therefore rises monotonically from the baseline as
the growth strengthens, so the reported shift is attributable entirely to the estimated
growth trend. Taking the no-growth count from the fleet's *forward* hazard (rather than
from the Crow-AMSAA since-inception average rate) is what removes the offset that a
purely cumulative-rate target would leave at $\beta_{g} = 1$.

## Limitations

The forecast projects the developmental improvement trend into service, and it inherits the
sampling behavior of the fleet's baseline Weibull fit. The Sensitivity Analysis (effect of
fleet assessment point) shows that when few failures have been observed the heavily censored
baseline fit is unstable, which propagates into the growth-adjusted forecast; a dependable
forecast therefore requires enough observed failures to pin down the fleet's baseline life
distribution. Results are
conditional on the growth trend continuing to be reflected in field reliability. If the
design is frozen at fielding and no further improvement occurs, the projection overstates
the gain and the no-growth baseline applies, so the forecast is best read as the optimistic
bound under continued improvement.

The present study demonstrates the mechanics and magnitude of the growth adjustment and
provides a reproducible framework for propagating developmental growth trends into
fleet-level forecasts. Establishing whether the method improves on the no-growth baseline in
predictive accuracy requires a validation study in which the fleet's true future life
genuinely continues along the estimated growth trend, so that the forecast can be compared
against a held-out ground truth. Such validation is left to future work.

Finally, the Monte Carlo analysis propagates estimation uncertainty in
the growth parameter $\beta_{g}$, though because the cumulative-data fit estimates are precise, that contribution is small and the spread is dominated by the conditional
simulation. Two uncertainty sources remain unrepresented: model-form uncertainty in the
choice of the Crow-AMSAA power law, and the input Weibull shape $\beta$, which is held
fixed at its fielded-fleet estimate across iterations rather than resampled. Holding
$\beta$ fixed is also a structural modeling choice, distinct from that sampling question:
reliability growth is treated as a scale effect that lengthens the characteristic life
$\eta$ while leaving the unit-level failure-mode shape $\beta$ unchanged. A design change
that *reshapes* the failure mode --- rather than delaying it --- would move $\beta$ in a
mode-specific direction that the growth parameter $\beta_{g}$ does not determine, and is
therefore outside the model's scope. Alternative growth models such as the piecewise NHPP
with change-point detection [@guo2010piecewise], available through segmented regression
[@muggeo2008segmented], may yield different growth counts and life distributions. The
simulation assumes a homogeneous fleet drawn from a single conditional Weibull
distribution and that failed units are not replaced; distinct subpopulations would
require separate analyses.

## Acknowledgements

The author thanks Anthropic's Claude Code for assistance with software development,
documentation, and manuscript preparation.

# References
